Poker Chips, Earthquakes, and Computation

### 2014-10-11 finaldraft.tex
finished simple model computation section

\begin{theorem}
Let $b$ be a board that becomes a stable board $b$ after $N$ earthquakes. Then for any other sequence of $N$ valid earthquakes the resultant board is identical to $b$. Furthermore the number of times that a square $x$ is targeted is invariant of the order in which the earthquakes occur. 
\end{theorem}
\begin{proof}
Given the original configuration of the board, let $A_i$ be the number of chips on square $i$. We know that square $i$ will be hit by at least 
\begin{equation*}
\lfloor \frac{A_i}{4} \rfloor
\end{equation*}
earthquakes in order to stabilize. Thus, somewhere in the sequence $\{E_i\}$, there exist $\lfloor \frac{A_i}{4} \rfloor$ earthquakes that target  square $i$. We note that these earthquakes will be valid in any position to the left of their original one because they are among the first $\lfloor \frac{A_i}{4} \rfloor$ to target a square that starts with $A_i$ chips on it. Thus using Lemma ~\ref{shiftlemma}, we can move these earthquakes to the front of the sequence without changing the final positioning of the board. We do this for all squares $i$.

@@

>>>>

\begin{theorem}
\label{thm:order}
Let $b$ be a board that becomes a stable board $b$ after $N$ earthquakes. Then for any other sequence of $N$ valid earthquakes the resultant board is identical to $b$. Furthermore the number of times that a square $x$ is targeted is invariant of the order in which the earthquakes occur. 
\end{theorem}
\begin{proof}
Given the original configuration of the board, let $A_i$ be the number of chips on square $i$. We know that square $i$ will be hit by at least 
\begin{equation*}
\lfloor \frac{A_i}{4} \rfloor
\end{equation*}
earthquakes in order to stabilize. Thus, somewhere in the sequence $\{E_i\}$, there exist $\lfloor \frac{A_i}{4} \rfloor$ earthquakes that target  square $i$. We note that these earthquakes will be valid in any position to the left of their original one because they are among the first $\lfloor \frac{A_i}{4} \rfloor$ to target a square that starts with $A_i$ chips on it. Thus using Lemma ~\ref{shiftlemma}, we can move these earthquakes to the front of the sequence without changing the final positioning of the board. We do this for all squares $i$.

@@

\begin{lemma}
Let $A(i)$ be the set of active nodes in an execution of board $b$ with input node $i$. Then $A(0) \subset A(1)$.
\end{lemma}

\begin{proof}
By Theorem~\ref{thm:order}, we can simulate an execution of the board with input be 0 before we apply earthquakes to the input node. We start with a board $b$ where the input is 1, so it has 4 chips. We simulate by applying earthquakes to the active nodes that are not the input. If the input node ever receives more than $8$ chips, we apply one earthquake to the input node. This way, we guarantee that the input node always contains more than $4$ chips. In addition, the input node does not contain more than $7$ chips for more than one consecutive earthquake.
We do this until every node except the input node has less than 4 chips.

Note that this execution is indistinguishable from the execution with input set to 0. This means that all nodes activated in the execution with input 0 are the same as the nodes activated in the above mentioned execution.

Once this is done, then we continue with the execution applying the earthquake to the input node. By Theorem~\ref{thm:order}, the final configuration will be the same as any execution of the board. So $A(0) \subset A(1)$.
\end{proof}

\subsection{More powerful model}
